On the analytic structure of the linear response susceptibility in open systems

### 2018-04-24 AnalyticStructure_main.tex
update

% This analysis reveals that the excitations spectrum is composed by two distinct sectors. The first one is realized by the (finite) set of localized excitation with associated discrete eigenvalues. 
% Extending the results of \cite{boffi2016} to the equation of motion of \eqref{ExcitationOperatorsDef3} we conclude that the elements of this category behave as \emph{observables} 
% quantities and the associated excitation energies have a true physical meaning, independently from the computational setup used for their representation. Conversely, the second sector is made by 
% a dense set of delocalized excitations with a continuum spectrum. The elements of this set behaves as the unbound virtual orbitals analyzed in \cite{boffi2016}....

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{tikzpicture}[scale=0.78]
% axes
\draw[->] (0,0) -- (8,0); 
\draw[->] (0,0) -- (0,6);
% orizontal lines
\draw[dashed,thick,color=gray] (0,5) -- (2,5);
\draw[very thick](2,5) -- (8,5);
\draw[dashed,thick,color=gray] (0,3) -- (4,3);
\draw[very thick] (4,3) -- (8,3);
\draw[dashed,thick,color=gray] (0,1) -- (6,1);
\draw[very thick] (6,1) -- (8,1);
% % vertical lines
\draw[dashed,very thin,color=black] (2,0) -- (2,5);
\draw[dashed,very thin,color=black] (4,0) -- (4,3);
\draw[dashed,very thin,color=black] (6,0) -- (6,1);
% % threshold levels labels
\node[left] at (0,5) {$p_h$};
\node[left] at (0,3) {$p_i$};
\node[left] at (0,1) {$p_1$};
\node[below] at (2,0) {$|\eps_h|$};
\node[below] at (4,0) {$|\eps_i|$};
\node[below] at (6,0) {$|\eps_1|$};
\node[below] at (4,-0.5) {excitation energy};
% bars at thresholds
%\foreach \Point in {(2,5),(4,3),(6,1)}{
%   \node at \Point {{\textbar}};%$\times$};
%}
\draw[very thick] (2,4.9) -- (2,5.1);
\draw[very thick] (4,2.9) -- (4,3.1);
\draw[very thick] (6,0.9) -- (6,1.1);
% dotted vertical lines
%\draw[dotted,thick,color=black] (-0.2,3.6) -- (-0.2,4.4);
%\draw[dotted,thick,color=black] (-0.2,1.6) -- (-0.2,2.4);
% discrete excitations
\foreach \Point in {(0.5,5),(1.0,5),(1.8,5),(0.8,3),(1.3,3),(1.8,3),(2.4,3),(2.8,3),(3.4,3),(3.8,3),(0.7,1),(1.5,1),(1.9,1),(2.6,1),(2.9,1),(3.5,1),(4.1,1),(4.7,1),(5.7,1)}{
    \node at \Point {\textbullet};
}
\end{tikzpicture}
\caption{\label{ExcitationLandscape} Excitation landscape. Excitation are split in horizontal line according to the threshold levels $p_m$. Filled bullet refers to localized
excitations whereas the tick black lines describe delocalized excitation with a continuum spectrum.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This analysis reveals that the excitations spectrum is composed by two distinct sectors. The first one is realized by the (finite) set of localized excitations with discrete eigenvalues. Extending 
the results achieved in \cite{boffi2016} for the KS virtual orbitals, to the equations for excited states \eqref{ExcitationOperatorsDef3}, we conclude that localized excitations behave as 
\emph{observable} quantities and the associated energies possess an absolute meaning, which should not depend from the computational setup used for their representation. 
The second sector contains the dense set of delocalized excitations with an associated continuum spectrum. These excitations behave as the unbound virtual orbitals analyzed in \cite{boffi2016} and 
are deeply affected by the computational setup used to represent them. In particular, the presence of a finite computational domain implies a fictitious discretization of the excitation levels and 
the associated spectrum exhibit an explicit dependence from the size of the domain. Excitations belonging to this sector thus lose their physical meaning and have to be only understood as a tool to 
express the spectral decomposition of the Liouvillian superoperator.

The typical structure of the excitation spectrum is depicted in figure \ref{ExcitationLandscape}. The plot evidences that below the first ionization energy $|\eps_h|$ only localized excitations contribute
to the spectrum. Conversely, in the intermediate energy interval that extends from $|\eps_h|$ to the absolute value of the deepest occupied orbital, the elements belonging to the discrete sector are embedded 
in a continuum of delocalized excitations. Lastly, for energies higher than $|\eps_1|$ only delocalized excitations are present.

@@

% Now we reformulate an analogous analysis looking at the susceptibility superoperator, implicit defined through \eqref{SusceptibilityDef1}. This approach leads to the concept of \emph{excitation 
% operators}, defined through a self-consistent operator eigenvalue equation. Excitations behave as singularity of the susceptibility operator and their knowledge provides a deeper insight 
% into the analytic structure of the linear response.

@@

We can express the fluctuation states in terms of the susceptibility functional
\be\lb{FluctuationStateExcitationDef1}
\ket{f_p(\omega)}   =
\left(\sum_E + \int\dd E   \right) 
\frac{\trace{\commutator{\dmnot}{\op{\tilde E}}\op\Phi(\omega)}}{\omega-E} \ket{\phi^E_p} \;. \nn
\ee 
This relation clarifies the role of the excitations in the spectral decomposition. We know from the analysis of section \ref{FluctuationState} that fluctuation states are localized
up to $\omega=|\eps_h|$.

@@

%\bibliographystyle{apsrev4-1}
\bibliography{Analytic_biblio}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{tikzpicture}[scale=0.78]
% axes
\draw[->] (0,0) -- (8,0); 
\draw[->] (0,0) -- (0,6);
% orizontal lines
\draw[dashed,thick,color=gray] (0,5) -- (2,5);
\draw[very thick](2,5) -- (8,5);
\draw[dashed,thick,color=gray] (0,3) -- (4,3);
\draw[very thick] (4,3) -- (8,3);
\draw[dashed,thick,color=gray] (0,1) -- (6,1);
\draw[very thick] (6,1) -- (8,1);
% % vertical lines
\draw[dashed,very thin,color=black] (2,0) -- (2,5);
\draw[dashed,very thin,color=black] (4,0) -- (4,3);
\draw[dashed,very thin,color=black] (6,0) -- (6,1);
% % threshold levels labels
\node[left] at (0,5) {$p_h$};
\node[left] at (0,3) {$p_i$};
\node[left] at (0,1) {$p_1$};
\node[below] at (2,0) {$|\eps_h|$};
\node[below] at (4,0) {$|\eps_i|$};
\node[below] at (6,0) {$|\eps_1|$};
\node[below] at (4,-0.5) {excitation energy};
% bars at thresholds
%\foreach \Point in {(2,5),(4,3),(6,1)}{
%   \node at \Point {{\textbar}};%$\times$};
%}
\draw[very thick] (2,4.9) -- (2,5.1);
\draw[very thick] (4,2.9) -- (4,3.1);
\draw[very thick] (6,0.9) -- (6,1.1);
% dotted vertical lines
%\draw[dotted,thick,color=black] (-0.2,3.6) -- (-0.2,4.4);
%\draw[dotted,thick,color=black] (-0.2,1.6) -- (-0.2,2.4);
% discrete excitations
\foreach \Point in {(0.5,5),(1.0,5),(1.8,5),(0.8,3),(1.3,3),(1.8,3),(2.4,3),(2.8,3),(3.4,3),(3.8,3),(0.7,1),(1.5,1),(1.9,1),(2.6,1),(2.9,1),(3.5,1),(4.1,1),(4.7,1),(5.7,1)}{
    \node at \Point {\textbullet};
}
\end{tikzpicture}
\caption{\label{ExcitationLandscape} Excitation landscape. Excitation are split in horizontal line according to the threshold levels $p_m$. Filled bullet refers to localized
excitations whereas the tick black lines describe delocalized excitation with a continuum spectrum.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This analysis reveals that the excitations spectrum is composed by two distinct sectors. The first one is realized by the (finite) set of localized excitations with discrete eigenvalues. Extending 
the results achieved in \cite{boffi2016} for the KS virtual orbitals, to the equations for excited states \eqref{ExcitationOperatorsDef3}, we conclude that localized excitations behave as 
\emph{observable} quantities and the associated energies possess an absolute meaning, which should not depend from the computational setup used for their representation.

Instead, the second sector contains the dense set of delocalized excitations with an associated continuum spectrum. These excitations behave as the unbound virtual orbitals analyzed in \cite{boffi2016} 
and are deeply affected by the computational setup used to represent them. In particular, the presence of a finite computational domain implies a fictitious discretization of the excitation levels and 
the associated spectrum exhibit an explicit dependence from the size of the domain. Excitations belonging to this sector thus lose their intrinsic physical meaning and have to be understood as a tool to 
express the spectral decomposition of the Liouvillian superoperator.

The typical structure of the excitations landscape is depicted in figure \ref{ExcitationLandscape}. The plot evidences that below the first ionization energy $|\eps_h|$ only localized excitations are 
present. Conversely, in the intermediate energy interval that extends from $|\eps_h|$ to the absolute value of the deepest occupied orbital, the elements belonging to the discrete sector are embedded 
in a continuum of delocalized excitations. Lastly, for energies higher than $|\eps_1|$ only delocalized excitations are present.

@@

%Excitations behave as singularity of the susceptibility operator and their knowledge provides a deeper insight into the analytic structure of the linear response.

@@

The analysis discussed above has evidenced that there exists both localized and delocalized excitations. While the role of the former is quite unambigous 
We can express the fluctuation states in terms of the susceptibility functional
\be\lb{FluctuationStateExcitationDef1}
\ket{f_p(\omega)}   =
\left(\sum_E + \int\dd E   \right) 
%\frac{\trace{\commutator{\dmnot}{\op{\tilde E}}\op\Phi(\omega)}}{\omega-E} \ket{\phi^E_p} \;. \nn
\frac{t_\Phi(E)}{\omega-E} \ket{\phi^E_p} \;,
\ee 
with $t_\Phi(E)=\trace{\commutator{\dmnot}{\op{\tilde E}}\op\Phi(\omega)}$.
This relation clarifies the role of the excitations in the spectral decomposition. We know from the analysis of section \ref{FluctuationState} that fluctuation states are localized
up to $\omega=|\eps_h|$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{apsrev4-1}
\bibliography{Analytic_biblio}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
